---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r}
require(plyr)
library(ggplot2)
data <- read.csv("activity.csv")
cleanData<-data[complete.cases(data),]
cdata <- ddply(cleanData, c("date"), summarise,sum    = sum(steps))
hist(cdata$sum,main="Histograms of Steps per day",xlab = "Number of steps per day")
```

## What is mean total number of steps taken per day?

### Mean and median
```{r}
mean<-mean(cdata$sum)
mean
median<-median(cdata$sum)
median
```

The mean of the steps per day is `r mean` and the median is `r median`


## What is the average daily activity pattern?
```{r}
stepsPerInterval <- ddply(cleanData, c("interval"), summarise,average    = mean(steps))
plot(stepsPerInterval$interval,stepsPerInterval$average,type="l",ylab="Average Steps Taken",xlab="Interval")
```
## What is interval with the max average number ofsteps
```{r}
interval <- stepsPerInterval[stepsPerInterval$average==max(stepsPerInterval$average),"interval"]
interval 
```
The interval with the maimum averagenumber of steps is `r interval`





## Imputing missing values

```{r}
#The total number of NA is 
nErrObs=nrow(data)-nrow(cleanData)
```
Total number of NAs is `r nErrObs`

###Description of the strategy 
For each NA i will find the average number of steps for that interval and substitute it

```{r}
dataSub<-data
for (i in 1:nrow(data)) {
  #if teh value is a NA find the average number in the interval and substitute it
  if(is.na(data$steps[i])){
    intervalI<-data[i,"interval"]
    meanIntervalI<-mean(stepsPerInterval[stepsPerInterval$interval==intervalI,"average"])
    dataSub[i,"steps"]<-meanIntervalI
  }
}
cdataSub <- ddply(dataSub, c("date"), summarise,sum    = sum(steps))
hist(cdataSub$sum,main="Histograms of Steps per day",xlab = "Number of steps per day")
```

### Mean and median
```{r}
meanSub<-mean(cdataSub$sum)
meanSub
medianSub<-median(cdataSub$sum)
medianSub
```

The mean of the steps per day of teh cleaned data is `r meanSub` and the median is `r medianSub`
The mean of the original data and the cleaned data should be the same since I'm adding the average
The difference between the means is `r meanSub-mean`  

## Are there differences in activity patterns between weekdays and weekends?

```{r}
#Convert the factor variable to a date
dataSub$date=as.POSIXlt(dataSub$date)
#compute the workday
dataSub$dayOfWeek=weekdays(dataSub$date)
#calculate the factors
dataSub$weekends = factor(weekdays(dataSub$date) %in% c("Sunday","Saturday"), labels = c("weekday","weekend"))

#create the data for teh weekday and weekend plots
aggregatedDataWeekday <- ddply(dataSub[dataSub$weekends=="weekday",], c("interval"), summarise,average    = mean(steps))
aggregatedDataWeekend <- ddply(dataSub[dataSub$weekends=="weekend",], c("interval"), summarise,average    = mean(steps))


par(mfrow = c(2,1))
#first subplot
plot(aggregatedDataWeekday$interval,aggregatedDataWeekday$average,type="l",main="Weekday",ylab="Avg number of steps",xlab="Interval")
#second subplot
plot(aggregatedDataWeekend$interval,aggregatedDataWeekend$average,type="l",main="Weekend",ylab="Avg number of steps",xlab="Interval")

```
The results indicate that we seem to be more active during the weekends  - probably since most of us sit at work. 



